---
title: "ReboundTools"
author: "Matthew Kuperus Heun"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{ReboundTools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.width = 5,
  fig.height = 3,
  fig.align = "center",
  fig.retina = 2
)
```

```{r setup, include = FALSE}
library(dplyr)
library(ReboundTools)
```


## Introduction

`ReboundTools` is an `R` package that provides functions to analyze 
energy rebound,  
the unanticipated reduction 
of the benefits of energy efficiency 
due to behavior change and economy-wide effects.
Many functions perform analysis calculations to move from 
known parameters to rebound estimates.
Graphing functions create rebound path graphs in 
energy, expenditure, and preference spaces.
Other graphing functions create sensitivity graphs. 
The functions in this package were used for the 
analyses and graphs in the paper **********.


## Getting started

The first step in a rebound analysis is
specifying parameters for the energy efficiency upgrade (EEU)
and rebound analysis, 
typically in a Excel spreadsheet.
The path to an example spreadsheet can be found with the function `sample_eeu_data_path()`.
Data can be loaded into a data frame as follows:

```{r}
eeu_data <- sample_eeu_data_path() %>% 
  load_eeu_data() # Uses the readxl package internally.
dplyr::glimpse(eeu_data)
```

Variable names (columns in the EEU spreadsheet)
are described by various constants in the package, 
including `eeu_base_params`, `key_analysis_vars`, and `rebound_stages`.
Get help on any of those constants to see descriptions of 
parameters, variables, and stages.

Rows of the EEU spreadsheet should describe the cases to be analyzed.
The example file at `sample_eeu_data_path()` contains two cases: 
a car and an electric lamp.
Columns of the EEU spreadsheet should contain rebound parameters for each case.
Required columns are shown in the example file and in the `eeu_data` data frame.

```{r}
colnames(eeu_data)
```

Rows can be used for sensitivity analysis, varying one or many parameters 
in each row of the spreadsheet.


## Performing the analysis

After data are loaded, rebound analyses can be performed.
A rebound analysis consists of calculating all rebound terms and intermediate results, 
once for each row of the EEU data frame.
A convenient function to perform the rebound analysis is named, aptly, `rebound_analysis()`.
Columns for intermediate and final results are added to the right of the data frame.

```{r}
rebound_data <- rebound_analysis(eeu_data)
dplyr::glimpse(rebound_data)
```

Note that `rebound_analysis()` is a convenience function that calls several 
helper functions in turn.
Each helper function calculates rebound parameters after one of the rebound effects.

```{r}
simple <- rebound_analysis(eeu_data)
complicated <- eeu_data %>% 
  calc_orig() %>%    # Calculate all parameters before the emplacement effect
  calc_star() %>%    # Calculate all parameters after the emplacement effect
  calc_hat() %>%     # Calculate all parameters after the substitution effect
  calc_bar() %>%     # Calculate all parameters after the income effect
  calc_tilde() %>%   # Calculate all parameters after the macro effect
  calc_Deltas() %>%  # Calculate all differences between stages
  calc_rebound()     # Calculate all rebound terms
all(simple == complicated)
```


## Rebound tables

`ReboundTools` contains two functions to create tables:
`stages_table()` and `rebound_results_table()`.


### Stages table

After a rebound analysis has been performed, 
the results can be displayed in a stages table, which has 
rebound parameters in rows and 
rebound stages in columns.

```{r, results="asis"}
digs <- matrix(c(rep(1, 7),  # eta_engr_units
                 rep(3, 7),  # eta
                 rep(3, 7),  # p_s
                 rep(0, 7),  # q_dot_s
                 rep(0, 7),  # E_dot_s
                 rep(0, 7),  # E_dot_emb
                 rep(0, 7),  # C_dot_s
                 rep(0, 7),  # C_dot_cap
                 rep(0, 7),  # C_dot_md
                 rep(0, 7),  # C_dot_o
                 rep(0, 7),  # N_dot
                 rep(0, 7)), # M_dot
               nrow = 12, ncol = 7, byrow = TRUE)
load_eeu_data() %>% 
  dplyr::filter(.data[[ReboundTools::eeu_base_params$case]] == "Car") %>% 
  dplyr::mutate(
    Case = NULL
  ) %>% 
  rebound_analysis() %>% 
  stages_table(digits = digs, 
               align = "lrrrrrr") %>% 
  print(type = "html",
        include.rownames = FALSE, 
        booktabs = TRUE,
        caption.placement = "top",
        sanitize.text.function = function(x) {x},
        format.args = list(big.mark = ","))
```


### Rebound results table

Rebound analysis results can also be displayed in a rebound results table.
A rebound results table shows rebound terms and values.

```{r, results="asis"}
load_eeu_data() %>% 
  dplyr::filter(Case == "Car") %>% 
  rebound_analysis() %>% 
  dplyr::mutate(
    Case = NULL
  ) %>%
  ReboundTools::rebound_results_table(include_subtotals = FALSE, 
                                      label = "tab:car_results", 
                                      digits = 1, 
                                      align = "rrr") %>% 
  print(type = "html",
        include.rownames = FALSE, 
        booktabs = TRUE,
        caption.placement = "top",
        hline.after = c(-1, 0, nrow(.)-1, nrow(.)),
        sanitize.text.function = function(x) {x},
        format.args = list(big.mark = ","))
```



## Rebound path graphs

Rebound graphs can be created with the `path_graphs()` function.
Three types of rebound graphs are available:

* Energy paths,
* Expenditure paths, and
* Preference paths.
   
The `graph_type` argument to `path_graphs()` controls which
type of path graph is returned
(one of "Energy", "Expenditure", or "Preferences").
By default, all three graph types are returned.
The `cases` argument determines which cases are plotted.
By default, all cases are returned.


### Energy path graph

An energy path graph shows the energy effects of the energy efficiency upgrade,
with direct energy consumption on the horizontal axis and
indirect energy consumption on the vertical axis.
Diagonal lines with negative slope indicate lines of constant energy consumption.
The thicker lines show 0 % rebound (lower left) and 100 % rebound (upper right).
Thinner lines show rebound after each effect in order from the 0 % line to the 100 % line:
emplacement, substitution, income, and macro.
The following example illustrates an energy path graph. 

```{r}
load_eeu_data() %>% 
  rebound_analysis() %>% 
  path_graphs(cases = "Car", graph_types = "Energy") +
  ggplot2::theme_classic()
```


### Expenditure path graph

The expenditure path graph shows expenditures for rebound effects. 
Direct expenditures for the energy service (including substitution and income takebacks)
are shown in the horizontal axis;
capital expenditures, maintenance and disposal expenditures, substitution effect savings, and
income effect expenditures are shown in the vertical axis. 
Diagonal lines with negative slope indicate lines of constant expenditure.
The lower-left line indicates expected expenditure options before 
capital expenditure changes, 
maintenance expenditure changes, or
behavior changes.
Note that the beginning (dot) and end (arrow) of the expenditure path both lie on the same
expenditure line, indicating that all freed cash 
(after the emplacement and substitution effects) 
is re-spent in the income effect.

```{r}
load_eeu_data() %>% 
  rebound_analysis() %>% 
path_graphs(cases = "Car", graph_types = "Expenditure") +
  ggplot2::theme_classic()
```


### Preferences path graph

The preferences path graph shows details of the substitution and income effects, 
with normalized energy service consumption on the horizontal axis
and normalized expenditures on other goods on the vertical axis.
Diagonal grid lines with negative slope indicate constant expenditure sum
of the energy service (horizontal axis) and other goods (vertical axis).
Negative-sloping lines show the tradeoff between consumption of the energy service 
(horizontal axis) and consumption of other goods (vertical axis).
The swooping, concave-upward grid lines indicate lines of constant utility:
indifference curves.

```{r warning = FALSE, message = FALSE}
load_eeu_data() %>% 
  rebound_analysis() %>% 
  path_graphs(cases = "Car", graph_types = "Preferences") +
  ggplot2::xlim(0.995, 1.06) + 
  ggplot2::ylim(0.995, 1.04) +
  ggplot2::theme_classic()
```


### Arguments to `path_graphs()`

In addition to the `graph_types` and `cases` arguments, 
the following arguments adjust the composition and appearance of graphs:

* `indexed`,
* `graph_types`, 
* `grid_types`, and
* `graph_params.`


#### `indexed`

Setting `indexed = TRUE` normalizes the rebound data to conditions
before emplacement of the energy efficient device, 
such that the starting point is always (1, 1).

```{r}
load_eeu_data() %>% 
  rebound_analysis() %>% 
  path_graphs(indexed = TRUE, cases = "Car", graph_types = "Energy") +
  ggplot2::theme_classic()
```

As discussed above, the `cases` argument tells which `Case`s should be plotted.
All cases are plotted by default, although the result is rarely pleasing.

```{r}
load_eeu_data() %>% 
  rebound_analysis() %>% 
  path_graphs(cases = c("Car", "Lamp"), 
                   graph_types = "Energy") +
  ggplot2::theme_classic()
```

When the cases are overlapping or very different (as above),
faceting and indexing may improve the appearance of the graph.

```{r, fig.width=5, fig.height=3, fig.align="center"}
load_eeu_data() %>% 
  rebound_analysis() %>% 
  path_graphs(indexed = TRUE,
                   graph_types = ReboundTools::graph_types$energy) +
    ggplot2::facet_wrap(facets = ReboundTools::eeu_base_params$case) +
  ggplot2::theme_classic()
```


#### `graph_types`

The `graph_types` argument tells which type of graph should be created. 
The default `ReboundTools::graph_types` returns all three graph types.
Again, faceting and indexing may be beneficial.

```{r}
load_eeu_data() %>% 
  rebound_analysis() %>% 
  path_graphs(indexed = TRUE, cases = "Car") +
  ggplot2::facet_wrap(facets = ReboundTools::graph_df_colnames$graph_type_col) +
  ggplot2::xlim(0.5, 1.2) +
  ggplot2::ylim(0.99, 1.08) +
  ggplot2::theme_classic()
```

Note that the `path_graphs()` function applies labels to the x and y axes
whenever the labels are unambiguous.
However, when multiple path graphs are included in the same figure, 
the x and y axis labels are unclear.
The user can apply labels appropriate to their situation.

```{r}
load_eeu_data() %>% 
  rebound_analysis() %>% 
  path_graphs(indexed = TRUE, cases = "Car") +
  ggplot2::facet_wrap(facets = ReboundTools::graph_df_colnames$graph_type_col) +
  ggplot2::xlim(0.5, 1.2) +
  ggplot2::ylim(0.99, 1.08) +
  ggplot2::xlab("my x label") + 
  ggplot2::ylab("my y label") +
  ggplot2::theme_classic()
```



```{r}
load_eeu_data() %>% 
  rebound_analysis() %>% 
  path_graphs(indexed = TRUE) +
  ggplot2::facet_grid(rows = vars(Case),
                      cols = vars(graph_type), 
                      scales = "free") +
  ggplot2::xlim(0.5, 1.2) +
  ggplot2::ylim(0.99, 1.08) +
  ggplot2::theme_classic()
```


#### `grid_types`

The `grid_types` argument tells which grids to include on the rebound graphs.
By default, grids are provided for all graph types.
In the example below, grids are provided only for the energy and expenditure path graphs.

```{r}
load_eeu_data() %>% 
  rebound_analysis() %>% 
  path_graphs(indexed = TRUE, grid_types = c("Energy", "Expenditure")) +
  ggplot2::facet_grid(rows = vars(Case),
                      cols = vars(graph_type), 
                      scales = "free") +
  ggplot2::xlim(0.5, 1.2) +
  ggplot2::ylim(0.99, 1.08) +
  ggplot2::theme_classic()
```


#### `graph_params`

While `indexed` controls _how_ the data are quantified and 
`graph_types` and `grid_types` control _what_ data are shown,
`graph_params` controls the _appearance_ of rebound graphs.
`graph_params` should be a named list.
The default `graph_params` list is
`path_graph_params`.

```{r}
ReboundTools::path_graph_params
```

Most parameters are self-explanatory.
For example, `ReboundTools::path_graph_params$dempl_colour` 
sets the colour of the direct emplacement path.

```{r}
my_graph_params <- path_graph_params
my_graph_params$dempl_colour <- "darkred"
load_eeu_data() %>% 
  rebound_analysis() %>% 
  path_graphs(indexed = TRUE,
              graph_types = ReboundTools::graph_types$energy, 
              graph_params = my_graph_params) +
  ggplot2::facet_wrap(facets = ReboundTools::eeu_base_params$case) +
  ggplot2::theme_classic()
```

See help for `path_graph_params` for additional information. 


## Sensitivity graphs

Sensitivity graphs can be created by the `sensitivity_graphs()` function.
A named list of sensitivity parameters is required.
These graphs are easiest described via examples.

The first example shows the sensitivity of total rebound ($Re_{tot}$) to 
the macro parameter ($k$).

```{r}
df <- load_eeu_data()
sens_params <- list(Car = list(k = seq(0.5, 1.5, by = 0.5)), 
                    Lamp = list(k = seq(0, 2, by = 1)))
sensitivity_graphs(rebound_data = df, parameterization = sens_params, 
                   x_var = "k", y_var = "Re_tot") +
  ggplot2::facet_wrap(facets = "Case", scales = "free_x") +
  ggplot2::scale_colour_manual(values = c(Re_tot = "black"), guide = FALSE) + 
  ggplot2::scale_size_manual(values = c(Re_tot = 0.5), guide = FALSE) + 
  ggplot2::scale_linetype_manual(values = c(Re_tot = "solid"), guide = FALSE) +
  ggplot2::labs(y = expression(Re[tot]), 
                colour = ggplot2::element_blank(),
                size = ggplot2::element_blank(),
                linetype = ggplot2::element_blank()) +
  ggplot2::theme_classic()
```

The next example shows multi-variate sensitivity.
Values of the macro parameter ($k$) are shown in rows of the lattice plot.
Uncompensated price elasticity of energy service consumption ($\epsilon_{\dot{q}_s,p_s,UC}$) 
is shown in columns of the lattice plot.
Total rebound ($Re_{tot}$) is given on the $y$-axis, and 
energy intensity of the economy ($I_E$) is given on the $x$-axis.
The cases (Car and Lamp) are shown as different lines.

```{r}
sens_params_2 <- list(Car = list(k = seq(0, 2, by = 0.5), 
                                 I_E = seq(2, 5, by = 1), 
                                 e_qs_ps_UC = seq(-0.5, -0.1, by = 0.1)), 
                      Lamp = list(k = seq(0, 2, by = 0.5),
                                  I_E = seq(2, 5, by = 1), 
                                  e_qs_ps_UC = seq(-0.5, -0.1, by = 0.1)))
# Choose which rebound variables to include and their order.
sensitivity_graphs(rebound_data = df, parameterization = sens_params_2, 
                   x_var = "I_E", 
                   y_var = "Re_tot",
                   line_var = "Case") +
  ggplot2::facet_grid(rows = ggplot2::vars(k), 
                      cols = ggplot2::vars(e_qs_ps_UC), scales = "free_y") +
  ggplot2::scale_colour_manual(values = c(Car = "black", Lamp = "red")) + 
  ggplot2::scale_size_manual(values = c(Car = 0.5, Lamp = 1.0)) + 
  ggplot2::scale_linetype_manual(values = c(Car = "solid", Lamp = "dashed")) + 
  ggplot2::labs(colour = ggplot2::element_blank(), 
                size = ggplot2::element_blank(),
                linetype = ggplot2::element_blank()) +
  ggplot2::theme_classic()
```


A third sensitivity example shows 

```{r}
sens_params_3 <- list(Car = list(eta_engr_units_star = seq(35, 50, by = 0.5)), 
                      Lamp = list(eta_engr_units_star = seq(70, 90, by = 5)))
# Choose rebound terms to include in the graph and their order
rebound_vars <- c("Re_dempl", "Re_emb", "Re_md", "Re_dsub", "Re_isub", 
                  "Re_dinc", "Re_iinc", "Re_macro")

sensitivity_graphs(rebound_data = df, parameterization = sens_params_3,
                   x_var = "eta_engr_units_tilde", 
                   y_var = rebound_vars) + 
  ggplot2::facet_wrap(facets = "Case", scales = "free_x") +
  ggplot2::scale_colour_manual(values = 
                                 c(Re_dempl = ReboundTools::path_graph_params$dempl_colour,
                                   Re_emb = ReboundTools::path_graph_params$emb_colour,
                                   Re_md = ReboundTools::path_graph_params$md_colour, 
                                   Re_dsub = ReboundTools::path_graph_params$dsub_colour,
                                   Re_isub = ReboundTools::path_graph_params$isub_colour, 
                                   Re_dinc = ReboundTools::path_graph_params$dinc_colour,
                                   Re_iinc = ReboundTools::path_graph_params$iinc_colour,
                                   Re_macro = ReboundTools::path_graph_params$macro_colour), 
                               breaks = rebound_vars) +
  ggplot2::scale_size_manual(values = 
                               c(Re_dempl = 0.2, 
                                 Re_emb = ReboundTools::path_graph_params$emb_size,
                                 Re_md = ReboundTools::path_graph_params$md_size, 
                                 Re_dsub = ReboundTools::path_graph_params$dsub_size,
                                 Re_isub = ReboundTools::path_graph_params$isub_size, 
                                 Re_dinc = ReboundTools::path_graph_params$dinc_size,
                                 Re_iinc = ReboundTools::path_graph_params$iinc_size,
                                 Re_macro = ReboundTools::path_graph_params$macro_size), 
                             breaks = rebound_vars) +
  ggplot2::scale_linetype_manual(values = 
                                   c(Re_dempl = ReboundTools::path_graph_params$dempl_linetype, 
                                     Re_emb = ReboundTools::path_graph_params$emb_linetype,
                                     Re_md = ReboundTools::path_graph_params$md_linetype,
                                     Re_dsub = ReboundTools::path_graph_params$dsub_linetype,
                                     Re_isub = "11",
                                     Re_dinc = ReboundTools::path_graph_params$dinc_linetype,
                                     Re_iinc = "11",
                                     Re_macro = ReboundTools::path_graph_params$macro_linetype), 
                                 breaks = rebound_vars) +
  ggplot2::labs(x = expression(tilde(eta)*" [mpg (Car) or lm/W (Lamp)]"), 
                y = "Re terms [-]", 
                colour = ggplot2::element_blank(),
                size = ggplot2::element_blank(),
                linetype = ggplot2::element_blank()) + 
  ggplot2::theme_classic()
```



## Conclusion

The `ReboundTools` package assists with analyzing rebound effects and 
displaying the results.
In particular, 

* the `rebound_analysis()` function performs rebound analysis 
on energy efficiency upgrades,
* the `stages_table()` function returns an `xtable` object
suitable for inclusion in papers and reports,
* the `path_graphs()` function returns `ggplot2` objects 
that show rebound path graphs, and 
* the `sensitivity_graphs()` function returns `ggplot2` objects
that show sensitivity analyses. 

Taken together, these functions provide important capabilities to analyze
rebound effects for energy efficiency upgrades.
