---
title: "ReboundTools"
author: "Matthew Kuperus Heun"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{ReboundTools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(dplyr)
library(ReboundTools)
```

## Introduction

`ReboundTools` is an `R` package that provides functions to analyze 
energy rebound,  
the unanticipated reduction 
of the benefits of energy efficiency 
due to behavior change and economy-wide effects.
Many functions perform analysis calculations to move from 
known parameters to rebound estimates.
Other functions create rebound graphs in 
energy, cost, and preference spaces.
The functions in this package were used for the 
analyses and graphs in the paper **********.


## Getting started

The first step is specifying parameters for the energy efficiency upgrade (EEU)
and rebound analysis, 
typically in a spreadsheet.
The path to an example spreadsheet can be found with the function `sample_eeu_data_path()`.
Data can be loaded into a data frame as follows:

```{r}
eeu_data <- sample_eeu_data_path() %>% 
  load_eeu_data()
dplyr::glimpse(eeu_data)
```

Variable names are described by various constants in the package:

```{r}
help("eeu_base_params")
help("key_analysis_vars")
help("rebound_stages")
```

Rows of the EEU spreadsheet should contain the cases to be analyzed.
The example file contains two cases: 
a car and an electric lamp.
Columns of the EEU spreadsheet should contain rebound parameters for each case.
Required columns are shown in the example file and `eeu_data` data frame.

```{r}
colnames(eeu_data)
```

Rows can be used for sensitivity analysis, varying one or many parameters in columns 
for each row of the spreadsheet.


## Performing the analysis

After data are loaded, rebound analyses can be performed.
A rebound analysis consists of calculating all rebound terms and intermediate results, 
once for each row of the EEU data frame.
A convenient function to perform the rebound analysis is `rebound_analysis()`.
Columns for intermediate and final results are added to the right of the data frame.

```{r}
rebound_data <- rebound_analysis(eeu_data)
dplyr::glimpse(rebound_data)
```

Note that `rebound_analysis()` is a convenience function that calls several 
helper functions in turn.
Each helper function calculates rebound parameters after one of the rebound effects.

```{r}
simple <- rebound_analysis(eeu_data)
complicated <- eeu_data %>% 
  calc_orig() %>%    # Calculate all parameters before the emplacement effect
  calc_star() %>%    # Calculate all parameters after the emplacement effect
  calc_hat() %>%     # Calculate all parameters after the substitution effect
  calc_bar() %>%     # Calculate all parameters after the income effect
  calc_tilde() %>%   # Calculate all parameters after the productivity effect
  calc_Deltas() %>%  # Calculate all differences between stages
  calc_rebound()     # Calculate all rebound terms
all(simple == complicated)
```


## Rebound graphs

Rebound path graphs can be created with the `rebound_graphs()` function.
Three rebound path graphs are available:

   * Energy paths,
   * Cost paths, and
   * Preference paths.

The following example illustrates the `rebound_graphs()` function.





